---
title: "Assignment B-1: Making a function"
author: "Berke Ucar"
date: "2023-10-30"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Assignment B-1: Making a function

In the mini data analysis project of STAT545A, I used a lot of string extraction tasks. For this assignment, I thought it would be fun and useful to gather those steps to a function. This is important since those kind of tasks (string extraction from a dataset) is a common practice in data science. 

One concern that I have is to allow this function to extract as many regular expression as possible from a single column, which is both flexible for the extraction and sturdy since extractions are done column by column generally with their specific extraction patterns. Since I do not perform the extraction for lots of columns, I believe this function is not too generalized (i.e. atomic to the column). The function adds a new column to the extracted patterns named <column_name>_<the_number_of_the_argument>. Since this function only requires a single column to be a string column, the function is not restrictive. 

As a design choice, I let columns that have NA values to be a proper input since they could possess a non NA value as well. I check this with counting the number of non NA values and checking if the count is equal to 0. Furthermore, it is possible to extract a pattern from a double column with ```str_extract``` method. So, I let the input column to be any kind of column as well. 

```{r}

#' Extract patterns from a column
#' 
#' @details 
#' Extracts the given regular expression(s) from the column of the tibble and makes a new column or new columns out of extraction(s). The column of extraction can be any type of column since str_extract allows extraction on every data type and returns a string as a result. This function returns the resulted tibble.
#' 
#' @param dataset tibble that holds the data for extraction. I believe "dataset" represents the input well since user needs to use a tibble that represents a dataset.
#' @param column base variable which indicates the column that the extraction will be performed on. The reason I named this "column" is that user can specify only one column and column is explanatory for that case
#' @param ... string(s) that hold(s) the regular expression pattern(s) for the extraction. Since user can specify different patterns for the column, I chose ellipsis for these parameters.
#'
#' @return dataset tibble with the addition of resulted columns from the extraction
extract_from_column <- function(dataset, column, ...){
  arguments <- c(...)
  
  # sanity checks
  if (length(arguments) == 0) stop("Sorry, you did not enter a pattern or patterns.")
  if (!is_tibble(dataset)) stop("Sorry, the dataset you passed is not a tibble, it is: ", class(dataset))
  if (nrow(dataset)==0) stop("Sorry, you entered an empty dataset.")
  # if (length(select(data, {{column}})) == 0) stop("Sorry, the dataset that you provided does not have a column")
  if (summarise(dataset, n=sum(!is.na({{column}})))[1] == 0) stop("Sorry, you entered a full NA column.")
  
  
  resulted_dataset <- dataset
  for (i in 1:length(arguments)){
    name <- paste("added_col_", i,sep="")
    resulted_dataset <- resulted_dataset %>% mutate( !!name := str_extract(pattern=arguments[i], string={{column}}))
  }
  
  return(resulted_dataset)
}
```

In order to demonstrate my function, I will use different datasets, columns, and patterns as well. Here it goes:

```{r}
library(datateachr)

print(extract_from_column(building_permits, address, "BC .*") %>% select(address, added_col_1)) # basic usage, extract a pattern from a string column. I only show the extracted column and the resource column here for the sake of the reader.
print(extract_from_column(steam_games, original_price, "\\d+\\.")%>%select(original_price, added_col_1)) # extraction of the decimal part of a double column, I only show the extracted column and the resource column here for the sake of the reader. !!! For this method since 0 is 0 in double as well, it cannot find the . hence it cannot extract the pattern
print(extract_from_column(steam_games, all_reviews, "\\(.*\\)", "\\d+%")%>%select(all_reviews, added_col_1, added_col_2)) # extraction of the multiple parts of a string column, I only show the extracted columns and the resource column here for the sake of the reader.
```
Now since I showed the general usage of the function, it is turn for the testing and edge case checking. The following cells demonstrate the testing. I test the following:

1. Without any patterns
2. A non-tibble dataset
3. An empty dataset
4. A column with all NA values 
5. A column that has both NA and non-NA values, this is not an edge case but to see if the function works properly or not

I know 3 and 4 are kind of the same but I would explain the difference as this: if the dataset is only declared but not initiated, it is empty. If the dataset has defined columns but no values or has NA values all across the colum, it is full NA column according to my definition.
```{r}
library(testthat)
na_column_steam_games <- steam_games %>% filter(is.na(mature_content))
df <- data.frame(col_1 = c("berke", "ucar", "bioinformatics"))
empty <-tibble(data.frame())


test_that( "Testing for the function extract_from_column", {
  expect_error(extract_from_column(steam_games, all_reviews), "Sorry, you did not enter a pattern or patterns.")
  expect_error(extract_from_column(df, col_1, ".*"), "Sorry, the dataset you passed is not a tibble, it is:")
  expect_error(extract_from_column(empty, col_1, ".*"), "Sorry, you entered an empty dataset.")
  expect_error(extract_from_column(na_column_steam_games, mature_content, ".*"), "Sorry, you entered a full NA column.")
})
```

